Some examples in extended logic counting classes

### 2015-09-22 tex/examples.tex
22 sep

Let $\L = \{P,N,\leq\}$, and $\A = \langle A, P^\A, N^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc DNF} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, $P$ and $N$ is a binary relations described as follows:
\begin{eqnarray*}
P^\A &=& \{(c,\ell)\mid \ell \mbox{ appears in the $c$-th clause in }\Psi\},\\
N^\A &=& \{(c,\ell)\mid \neg \ell \mbox{ appears in the $c$-th clause in }\Psi\}.
\end{eqnarray*}

@@

\subsection{{\sc \#HORN-SAT}}
We will now show that {\sc \#DNF} $\in\EtwoE$.\\

>>>>

Let $\L = \{P,N,\leq\}$, and $\A = \langle A, P^\A, N^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc DNF} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, $P$ and $N$ is a binary relations described as follows:
\begin{eqnarray*}
P^\A &=& \{(c,\ell)\mid \ell \mbox{ appears in the $c$-th clause of }\Psi\},\\
N^\A &=& \{(c,\ell)\mid \neg \ell \mbox{ appears in the $c$-th clause of }\Psi\}.
\end{eqnarray*}

@@

\subsection{{\sc \#HORN-SAT}}
We will now show that {\sc \#HORN-SAT} $\in\EoneU$.\\

### 2015-09-22 tex/s-one-fo.tex
22 sep

>>>>

\begin{theo}
	{\sc \#DNF} is hard for $\E1$ under parsimonious reductions. 
\end{theo}
\begin{proof}
	Let $f$ be an arbitrary function in $\E1$ and let $f_{\textsc{DNF}}$ be the function that characterizes {\sc \#DNF}. We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{DNF}}(h(\A))$, and which can be computed in polynomial time. Let $\psi$ be such that $f = f_{\psi}$. We separate the proof in three cases:
	\begin{enumerate}
		\item All free variables in $\psi$ are of first order. Let $\psi = \exists x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = f(\A)$. First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$. Let $n$ be the number of such assignments to $\z$. Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$. The DNF formula we return is
		\[
		\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
		\]
		which has exactly $n$ possible assignments.
		\item There is only second-order free variables in $\psi$. Let $\psi = \exists x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a DNF-formula $\Phi = f(\A)$ as follows:

		First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$. Let each $\varphi_{i}(\x,\X)$ be of the form
		\[
		\varphi_{i}(\x,\X) = X_{s_1}(\x_{s_1}) \wedge \cdots \wedge X_{s_m}(\x_{s_m}) \wedge \neg X_{t_1}(\x_{t_1}) \wedge \cdots \wedge \neg X_{t_n}(\x_{t_n}) \wedge \alpha_i(\x),
		\]
		where every $\x_{s_i}$ and $\x_{t_j}$ have the corresponding number of variables, and $\alpha(\x)$ is a FO-formula.

		Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha(\a)\}$. There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time. We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in \bigcup_{j = 1}^c A^j$, and for each $\ell\in\{1,\ldots,r\}$. Let $\p$ be the tuple of all such variables. Also, let $\rho_{\ell}(\a)$ be the tuple that results of reordering $\a$ using the reordering of the variables of $\x$ in $\x_{\ell}$.

		Finally, for each $\varphi_i(\x,\X)$ the DNF formula we return is
		\[
		\Phi_i = \bigvee_{\a\in\Gamma_i} p^{s_1}_{\rho_{s_1}(\a)} \wedge \cdots \wedge p^{s_m}_{\rho_{s_m}(\a)} \wedge \neg p^{t_1}_{\rho_{t_1}(\a)} \wedge \cdots \wedge \neg p^{t_n}_{\rho_{t_n}(\a)}.
		\]
		and $\Phi = \bigvee_{i = 1}^k \Phi_i$.

		Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$. We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
		\item Let $\psi = \exists x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\X = (X_1,\ldots,X_r)$ and $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a DNF-formula $\Phi = f(\A)$ as follows:

		As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$. Let each $\varphi_{i}(\x,\X,\z)$ be of the form
		\[
		\varphi_{i}(\x,\X,\z) = X_{s_1}(\x_{s_1},\z_{s_1}) \wedge \cdots \wedge X_{s_m}(\x_{s_m},\z_{s_m}) \wedge \neg X_{t_1}(\x_{t_1},\z_{t_1}) \wedge \cdots \wedge \neg X_{t_n}(\x_{t_n},\z_{t_1}) \wedge \alpha_i(\x,\z),
		\]
		where every $\x_{s_i}$, $\z_{s_i}$, $\x_{t_j}$ and $\z_{t_j}$ have the corresponding number of variables, and $\alpha(\x)$ is a FO-formula.

		We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\b\in \bigcup_{j = 1}^{c+d} A^j$, for each $\e\in A^d$, and for each $\ell\in\{1,\ldots,r\}$. Let $\p$ be the tuple of all such variables. Also, let $\rho_{\ell}(\a)$ be the tuple that results of reordering $\a$ using the reordering of the variables of $(\x,\z)$ in $(\x_{\ell},\z_{\ell})$.

		For each $\e\in A^d$, we follow the following procedure. Eor each $\varphi_i(\x,\X)$, we generate the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
		\[
		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{s_1,\rho_{s_1}(\a,\e)} \wedge \cdots \wedge p^{\e}_{s_m,\rho_{s_m}(\a,\e)} \wedge \neg p^{\e}_{t_1,\rho_{t_1}(\a,\e)} \wedge \cdots \wedge \neg p^{\e}_{t_n,\rho_{t_n}(\a,\e)} \wedge \beta_{\e},
		\]
		where $\beta_{\e}$ is defined as
		\[
		\beta_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e}} \bigwedge_\bigvee_{\b\in A^(c+d)} \bigwedge_{\ell = 1}^r p^{\e}_{\ell,\b}.	
		\]
	\end{enumerate}
\end{proof}
